---
title: "Data Wrangling Exercise 2"
author: Anuradha Pani
output: html_notebook
---
***
### Prerequisites
Install all required packages and load them to R 
```{r}
library(tidyverse)
library(data.table)
```

0: Load Titanic data into R  
All input and output data/CSV files are in a dircetory named "data", in the current working dircetory.
```{r}
titanic <- fread("data/titanic_original.csv")
head(titanic)
tail(titanic)
```
  
1: Port of Embarkation
```{r}
# Port of embarkation
# Check to see how many missing values are there before correction
#sum(titanic$embarked == "" | is.na(titanic$embarked))

# Replace missing values (NA or empty strings) with "S"
titanic$embarked[titanic$embarked == "" | is.na(titanic$embarked)] <- "S"

# Check to see how many missing values are there after correction: should be zero.
#sum(titanic$embarked == "" | is.na(titanic$embarked))

```

2: Age
```{r}
# Age 1. replace missing values with mean age
titanic$age[is.na(titanic$age)] <- mean(titanic$age, na.rm = TRUE)
```

Age: Other ways to replace missing age
We can adopt the following two strategies:  
1. Find the age, which has the largest count and use that  
```{r}
titanic %>% 
  group_by(age) %>% 
  summarise(cnt = n()) %>% 
  arrange(desc(cnt))
```  
2. Use the median value  
```{r}
#titanic$age[is.na(titanic$age)] <- median(titanic$age, na.rm = TRUE)
```
  
3. Lifeboat
```{r}
# Fill missing values with a dummy string
titanic$boat[is.na(titanic$boat) | titanic$boat == ""] <- "NA"
```
  
    
4. Cabin
```{r}
titanic <- titanic %>% 
        mutate(has_cabin_number = ifelse(titanic$cabin == "", 0,1))

# Display new column to verify
titanic %>% 
        select(cabin, has_cabin_number) %>% 
        head(15)
```

Write the final data table/data frame to a CSV, "titanic_cleaned.csv", in the data directory
```{r}
write_csv(titanic, "data/titanic_cleaned.csv")

# Display the final clean data table
head(titanic, 25)
```
### End
***
