---
title: "Data Wrangling Exercise 1"
author: "Anuradha Pani"
output: html_notebook
---

### Prerequisites
Install all required packages and load them to R      
```{r}
# Packages required:
library(tidyverse)
```
      
0: Load the CSV into an R dataframe.  
All CSV files are in a directory called 'data', which is in the current working directory
```{r}
# 0: Load data in RStudio
# NOTE: 
brands <- suppressMessages(read_csv("data/refine_original.csv"))
head(brands, 15)
```
  1: Clean brand names  
  
```{r}
# 1: Clean up brand names
# Change all case to lower first
brands$company <- tolower(brands$company)

# Use regular expression to clean brad names
# philips
pattern_phil = "^[pf].*ps$"
replace_phil = "philips"
brands$company <- gsub(pattern_phil, replace_phil, brands$company)

# akzo
pattern_akzo = "^[ak].*[o0]$"
replace_akzo = "akzo"
brands$company <- gsub(pattern_akzo, replace_akzo, brands$company)

# van houten
pattern_vh = "^van.*ten$"
replace_vh = "van houten"
brands$company <- gsub(pattern_phil, replace_phil, brands$company)

# unilever
pattern_uni = "^uni.*"
replace_uni = "unilever"
brands$company <- gsub(pattern_uni, replace_uni, brands$company)
```

2: Separate product code and number
First, we'll rename column for Product code and number, so that it can be used easily with tidyr functions.
```{r}
# 2: Separate product code and number
# Rename second column
colnames(brands)[2] <- "product_code_num"

# Will use the separate() function from tidyr
brands <- separate(brands, product_code_num, c("product_code", "product_number"), sep = "-")
```

3: Add product categories
```{r}
# Add a column for product categories
brands <-brands %>% 
        mutate(product_category =
                       ifelse(product_code == "p", "Smartphone", 
                              ifelse(product_code == "q", "Tablet", 
                                     ifelse(product_code == "v", "TV", 
                                            ifelse(product_code == "x", "Laptop", NA)))))


```

4: Add Full Address for Geocoding
Combine the address, city and country columns separated by a comma
```{r}
# 4: Add full address for geocoding
brands <- brands %>% 
        mutate(full_address = paste(address, city, country, sep = ","))        
```

5: Create dummy variable for product category and company name
```{r}
# 5: Create dummy variable for product category and company name
## Dummy company names
brands <- brands %>% 
        mutate(company_philips = ifelse(company == "philips", 1, 0)) %>% 
        mutate(company_akzo = ifelse(company == "akzo", 1, 0)) %>% 
        mutate(company_van_houten = ifelse(company == "van_houten", 1, 0)) %>% 
        mutate(company_unilever = ifelse(company == "unilever", 1, 0))


## Dummy product categories
brands <- brands %>% 
        mutate(product_smartphone = ifelse(product_code == "p", 1, 0)) %>% 
        mutate(product_tv = ifelse(product_code == "v", 1, 0)) %>% 
        mutate(product_laptop = ifelse(product_code == "x", 1, 0)) %>% 
        mutate(product_tablet = ifelse(product_code == "q", 1, 0))
```

Write final dataframe to "refine_clean.csv"
```{r}
write_csv(brands, "data/refine_clean.csv")

# Display final clean dataframe
head(brands, 15)
```